---
title: "CCBR-668"
author: "Maggie Cam"
date: "December 19, 2015"
output: html_document
---

## *CCBR-668*
### *Maggie Cam*
### *Dec 16, 2015*

## **Background**
In Vitro Drug Screening in cells grown on different substrates 

Goal: To find genes differentially expressed in cells grown on substrates of different hardness (glass, medium, low)

## **Data Processing**

```{r setup, echo=FALSE, warning=FALSE}

#source("https://bioconductor.org/biocLite.R")
#biocLite("arrayQualityMetrics")
#biocLite("AnnotationForge")
#biocLite("AnnotationDbi")
#biocLite("human.db0")
#install.packages('knitr', repos='http://www.rforge.net/', type='source')

#Download annotations
#download.file("http://www.affymetrix.com/Auth/analysis/downloads/na35/ivt/PrimeView.na35.annot.csv.zip", "PrimeView.na35.annot.csv.zip")
#unzip("PrimeView.na35.annot.csv.zip")

library(primeviewcdf)
library(affy)
library(limma)
library(genefilter)
library(AnnotationForge)
library(AnnotationDbi)
library(human.db0)

#makeDBPackage("HUMANCHIP_DB", affy = TRUE, prefix="primeview",
#              fileName="PrimeView.na35.annot.csv/PrimeView.na35.annot.csv", 
#              baseMapType="gbNRef",outputDir=".",
#              version = "12.1.0", manufacturer="Affymetrix",
#              manufacturerUrl="http://www.affymetrix.com")
#install.packages("primeview.db", repos=NULL, type = "source")

library(primeview.db)
library(knitr)
library(rgl)

knit_hooks$set(webgl = hook_webgl)
knit_hooks$set(rgl = function(before, options, envir) {
  if (!before) {
    ## after a chunk has been evaluated
    if (rgl.cur() == 0) return()  # no active device
    name = paste(options$fig.path, options$label, sep = '')
    rgl.snapshot(paste(name, '.png', sep = ''), fmt = 'png')
    return(paste('\\includegraphics{', name, '}\n', sep = ''))
  }
})

knit_hooks$set(webgl = hook_webgl)
```
arrayQualityMetrics:::intgroupColors
function (x) 
{
    if (length(x$intgroup) > 0) {
        colors = c(brewer.pal(9, "Set1"), brewer.pal(8, "Dark2"))
        fac = as.factor(x$pData[[x$intgroup[1]]])
        fac = maximumLevels(fac, n = length(colors))
        colors = colors[seq_len(nlevels(fac))]
        ac = colors[as.integer(fac)]
        key = list(rect = list(col = colors), text = list(levels(fac)), 
            rep = FALSE)
    }
    else {
        key = NULL
        ac = rep("#1F78B4", x$numArrays)
    }
    list(arrayColors = ac, key = key)
}
```

####*Load data from local directory:*

```{r,eval=TRUE, echo=FALSE}

setwd("/Users/maggiec/GitHub/Maggie/ccbr668/data/")

celFiles=read.table("covdesc",header=TRUE)[,1]
pd<-read.AnnotatedDataFrame("covdesc",header=TRUE,sep="\t")
x <- varMetadata(pd)
x <- data.frame(x, channel = "_ALL_")
varMetadata(pd) <- x

affyData = ReadAffy()
phenoData(affyData) = pd

colnames(affyData)
gns <- select(primeview.db, featureNames(affyData),
              c("SYMBOL","GENENAME"))

####*QC Check: Look at raw signal distribution and median expression levels*
```

```{r, echo=FALSE,webgl=TRUE}
library(ggplot2)
library(reshape)
library(plotly)
library(matrixStats)

e=exprs(affyData)
df.m <- melt(as.data.frame(e))
ggplot(df.m) + 
  geom_density(aes(x = value, colour = variable)) + labs(x = NULL) +
  theme(legend.position='right') + scale_x_log10() + ggtitle("Raw Counts")

par(mar=c(10,7,1,1))
boxplot(log(value)~variable,las=2,data=df.m,main="Raw Signal", 
  	ylab="Signal",col=as.numeric(as.factor(pData(affyData)$Sample)))

#Plots for Normalized Data

eset = rma(affyData)
e <- exprs(eset)
df.m <- melt(as.data.frame(e))
ggplot(df.m) + 
  geom_density(aes(x = value, colour = variable)) + labs(x = NULL) +
  theme(legend.position='right') + scale_x_log10() + ggtitle("Raw Counts")

par(mar=c(10,7,1,1))
boxplot(log(value)~variable,las=2,data=df.m,main="Raw Signal", 
  	ylab="Signal",col=as.numeric(as.factor(pData(eset)$Sample)))

#Heatmap for normalized data
colorRange = rgb(seq(0, 1, l=256),
                   seq(0, 1, l=256),
                   seq(1, 0, l=256))
te=t(e) 
hc=hclust(dist(te))
plot(hc)


reorderfun = function(d,w) { d }
png("heatmap.png", res=150, height=22,width=17,units="in")

e <- exprs(eset)
te <- t(e)
tedf = as.data.frame(te)
tedf=cbind(pData(eset),tedf)

colnum=dim(tedf)[2]
tedf.mat=(as.matrix(tedf[, 4:colnum]))
pca=prcomp(tedf.mat)
Treat <- factor(tedf$Substrate) 

library(rgl)
open3d() 

plot3d(pca$x[,1:3],col=as.numeric(as.factor(Treat)), 
       type="s",size=2)
group.v<-as.vector(paste(tedf$Substrate,tedf$Rep))
text3d(pca$x, pca$y, pca$z, group.v, cex=0.6, adj = 1.5) 
rgl.postscript("pca3d_Qindiv.pdf","pdf")

```
####Run Differential Expression analysis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
design <- model.matrix(~0+Treat)

fit <- lmFit(eset, design)
cm <- makeContrasts(LowvsGlass = TreatLow-TreatGLASS,
                    MedvsGlass = TreatMed-TreatGLASS,
                    MedvsLow = TreatMed-TreatLow,
                    levels=design) 
fit2 <- contrasts.fit(fit, cm)
fit2 <- eBayes(fit2)
fit2$genes <- gns[!duplicated(gns[,1]),]
fit2$genename <- gns[!duplicated(gns[,2]),]
topTable(fit2, coef=3)
topTable(fit2, coef=2)
finalres=topTable(fit2,sort="none",n=Inf)

FC = 2^fit2$coefficients
FC = ifelse(FC<1,-1/FC,FC)
colnames(FC)=paste(colnames(FC),"FC",sep="_")
colnames(FC)
pvalall=fit2$p.value
colnames(pvalall)=paste(colnames(pvalall),"pval",sep="_")
colnames(pvalall)
topgenes=topTable(fit2, coef="LowvsGlass",n=50,sort.by="p")

LowvsGlass=p.adjust(fit2$p.value[,1], method='BH')
MedvsGlass=p.adjust(fit2$p.value[,2], method='BH')
MedvsLow=p.adjust(fit2$p.value[,3], method='BH')
pvaladjall=cbind(LowvsGlass,MedvsGlass,MedvsLow)
colnames(pvaladjall)=paste(colnames(pvaladjall),"adjpval",sep="_")

finalres = as.data.frame(cbind(finalres[,1:3],e,FC,pvalall,pvaladjall))
head(finalres)


#write.table(finalres,file="All_Results_Substrates.txt",          row.names=FALSE,sep="\t",quote=FALSE)
```

```{r, echo=FALSE, message=FALSE,warning=FALSE}

library(d3heatmap)
library(reshape2)

#Original Heatmap (notice batch effect)
topgenes_data=eset[rownames(eset) %in% rownames(topgenes),]
colnames(topgenes_data)=colnames(eset)
topgenes_data=topgenes_data[match(rownames(topgenes),rownames(topgenes_data)),]
rownames(topgenes_data) = topgenes$SYMBOL
d3heatmap(topgenes_data, scale = "row", dendrogram = "none", main = "title",
    color = "YlOrRd",cexRow=1,main="topgenes")
```
