
---
title: "RNAseq in embryonic skin of Shh, fgf20, edar mutant mice"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K128m",
      "-RTS"
    ]
---
## *CCBR-620*
### *Maggie Cam*
### *Oct 15, 2015*

## **Background**
To detect different gene expression in a small subpopulation (<1%) of skin, call Merkel cell, upon shh, fgf20, and edar mutation. Aim to analyze the connection between the three different pathways, and try to get clue of the genes participating in Merkel cell development.

E15.5 embryo skin from 2 mutant lines:
 
edar mutant = mutation in gene ectodysplasin-A receptor, Single point mutation (Mutation details: This allele involves a G to A transition mutation at nucleotide 1,135 that causes the amino acid change: glutamate to lysine at position 379 (E379K). (J:56496))  
                  phenotype (http://www.informatics.jax.org/allele/allgenoviews/MGI:1856018) = abnormal coat/hair morphology, darkened coat color  

fgf20 mutant = fgf20-β-galactosidase knock-in allele
                 phenotype= no guard hair in adult mouse back skin

## **Data Processing**

```{r setup, echo=TRUE, warning=FALSE}

source("http://bioconductor.org/biocLite.R")
#biocLite("edgeR")
library(Rsubread)
library(limma)
library(edgeR)
library(ggplot2)
library(rgl)
library(knitr)


knit_hooks$set(rgl = function(before, options, envir) {
  if (!before) {
    ## after a chunk has been evaluated
    if (rgl.cur() == 0) return()  # no active device
    name = paste(options$fig.path, options$label, sep = '')
    rgl.snapshot(paste(name, '.png', sep = ''), fmt = 'png')
    return(paste('\\includegraphics{', name, '}\n', sep = ''))
  }
})

knit_hooks$set(webgl = hook_webgl)
```
####*This part of the program was run on biowulf, and data transferred to a local directory*

```{r, eval=FALSE}
#Command line version
module load subread
x=$(ls *.bam)
featureCounts -p -T 8 -s 2 -p -t exon -g gene_id -a /data/maggiec/RNASeq/Genomes/mm10/gencode.vM4.all.gtf -o counts_ss.txt $x

#Used R version:
gtf="data/maggiec/RNASeq/Genomes/mm10/gencode.vM4.all.gtf"
targets <- readTargets()
fc <- featureCounts(files=targets$bam,isGTFAnnotationFile=TRUE,nthreads=32,
      annot.ext=gtf,GTF.attrType="gene_name",strandSpecific=2,isPairedEnd=TRUE)
x <- DGEList(counts=fc$counts, genes=fc$annotation)

```
####*Load data from local directory:*

```{r,eval=TRUE}
setwd("/Users/maggiec/GitHub/Maggie/ccbr620/")
#load("data/ssData.RData")
#fc=fc_ss
load("data/RNASeq_orig.RData")
mapped=read.delim(file = "data/mapped.txt",header = TRUE)
targets$mapped=mapped$mapped
targets$Phenotype[c(7,8,10)] <- "fgf20lz/lz mutant"
targets$Phenotype[c(9,11,12)] <- "fgf20lz/+ control"

ls()
targets
fc$stat

fc1=mat=fc$counts
tfc1=t(fc1)
filter <- apply(fc1, 1, function(x) length(x[x>5])>=1)
fc1filt <- fc1[filter,]
genes <- rownames(fc1filt)

```

####*Mapping Rate:*

```{r, fig.width=10, fig.height=5, warning=FALSE,echo=TRUE}
options(scipen=9)
map=as.numeric(targets$mapped)
names(map)=sapply(strsplit(targets$bam,split="\\."),function(x) x[1])
barplot(map, col=as.numeric(as.factor(targets$Phenotype)),las = 2,cex.axis = 0.7)
```


####*QC Check: Look at raw signal distribution and median expression levels*

```{r fig.width=10, fig.height=5, echo=TRUE, warning=FALSE,message=FALSE}

library(ggplot2)
library(reshape)

fc1filt=as.data.frame(fc1filt)
fc1filtnames=sapply(strsplit(colnames(fc1filt),split="_"),function(x) x[2])
fc1filtnames=sapply(strsplit(fc1filtnames,split="\\."),function(x) x[1])
colnames(fc1filt)=paste(fc1filtnames,targets$Phenotype)

df.m <- melt(as.data.frame(fc1filt))
ggplot(df.m) + 
  geom_density(aes(x = value, colour = variable)) + labs(x = NULL) +
  theme(legend.position='right') + scale_x_log10() + ggtitle("Raw Counts")

par(mar=c(10,7,1,1))
boxplot(log(value)~variable,las=2,data=df.m,main="Raw Signal", 
  	ylab="Counts",col=as.numeric(as.factor(targets$Phenotype)))

```

####*Data is normalized by TMM and quantile normalization*

```{r,fig.width=10, fig.height=5, echo=TRUE, warning=FALSE,message=FALSE}

x <- DGEList(counts=fc$counts, genes=fc$annotation)
isexpr <- rowSums(cpm(x)>0.5) >= 2
hasannot <- rowSums(is.na(x$genes))==0
x <- x[isexpr & hasannot,keep.lib.sizes=FALSE]

#Number of filtered genes 
dim(x)

x <- calcNormFactors(x)

par(mfrow=c(2,6))
for (i in 1:dim(x)[2]){
plotMD(cpm(x, log=TRUE), column=i)
abline(h=0, col="red", lty=2, lwd=2)
}

```

####*Run Voom*
```{r}
#Do analysis for entire group
celltype <- factor(targets$Phenotype)
design <- model.matrix(~celltype)
design
v <- voom(x,design,plot=TRUE,normalize="quantile")
```

####*After Normalization*

```{r, echo=TRUE,webgl=TRUE}

colnames(v$E)=paste(fc1filtnames,targets$Phenotype)
df.m <- melt(as.data.frame(v$E))

ggplot(df.m) + 
  geom_density(aes(x = value, colour = variable)) + labs(x = NULL) +
  theme(legend.position='right') + ggtitle("Normalized Counts")

par(mar=c(10,7,1,1))
boxplot(value~variable,las=2,data=df.m,main="Normalized Signal", 
  	ylab="Counts",col=as.numeric(as.factor(targets$Phenotype)))

edf=as.matrix(v$E)
tedf= t(edf)
pca=prcomp(tedf,scale.=T)
tedf1 = data.frame(tedf)
Phenotype=sapply(strsplit(targets$Phenotype,split=" "),function(x) x[1])
cell_rep=paste(Phenotype,targets$Batch,sep=".")
tedf1$group = as.factor(Phenotype)

plot(pca,type="lines")  #Decide how many PC's are relevant for plotting
pca$x[,1:3]  #look at first 3 PC's

plot3d(pca$x[,1:3],col = as.integer(tedf1$group),type="s",size=2)
group.v<-as.vector(cell_rep)
text3d(pca$x, pca$y, pca$z, group.v, cex=1.0, adj = 1.2) 
#rgl.postscript("pca3d_indiv.pdf","pdf")
```

####*Preprocessing of edar group (voom): 2 batches of pups (voom)*

####*Design: model.matrix(~Batch+celltype)*####

```{r, echo=TRUE, webgl=TRUE}
targets[1:6,]
celltype <- factor(targets$Phenotype[1:6])
Batch <- factor(targets$Batch[1:6])
cell_rep=paste(celltype,Batch,sep=".")
design1 <- model.matrix(~Batch+celltype)
design1
v1 <- voom(x[,1:6],design1,plot=TRUE,normalize="quantile")

edf=as.matrix(v$E[,1:6])
tedf= t(edf)
pca=prcomp(tedf,scale.=T)
tedf1 = data.frame(tedf)

Phenotype=sapply(strsplit(targets$Phenotype[1:6],split=" "),function(x) x[1])
Batch=targets$Batch[1:6]
tedf1$group = as.factor(Phenotype)
plot(pca,type="lines")  #Decide how many PC's are relevant for plotting
pca$x[,1:3]  #look at first 3 PC's

plot3d(pca$x[,1:3],col = as.integer(tedf1$group),type="s",size=2)
group.v<-as.vector(paste(Phenotype,Batch))
text3d(pca$x, pca$y, pca$z, group.v, cex=1.0, adj = 1.2) 
#rgl.postscript("pca3d_indiv_edar_batch.pdf","pdf")

```

####*Preprocessing of Fgf group (voom): 1 batch of pups*

####*Design: model.matrix(~celltype)*####

```{r, echo=TRUE, webgl=TRUE}
#Process the Fgf group
targets[7:12,]
celltype <- factor(targets$Phenotype[7:12])
design2 <- model.matrix(~celltype)
design2
v2 <- voom(x[,7:12],design2,plot=TRUE,normalize="quantile")

#fgf:
edf=as.matrix(v$E[,7:12])
tedf= t(edf)
pca=prcomp(tedf,scale.=T)
tedf1 = data.frame(tedf)

Phenotype=sapply(strsplit(targets$Phenotype[7:12],split=" "),function(x) x[1])
tedf1$group = as.factor(Phenotype)
plot(pca,type="lines")  #Decide how many PC's are relevant for plotting
pca$x[,1:3]  #look at first 3 PC's

plot3d(pca$x[,1:3],col = as.integer(tedf1$group),type="s",size=2)
group.v<-as.vector(Phenotype)
text3d(pca$x, pca$y, pca$z, group.v, cex=1.0, adj = 1.2) 
#rgl.postscript("pca3d_indiv_fgf.pdf","pdf")
```

####*Genes of Interest*

```{r, fig.width=10, fig.height=5, warning=FALSE,echo=TRUE}

genes=c("Sox2","Atoh1","Shh","Gli1","Edar","Traf6","Tnfrsf19","Lta","Ltb","Nfkb1")
vdf=cbind(v1$E,v2$E)
vdf=vdf[,c(1,2,3,4,5,6,9,11,12,7,8,10)]
genes_count=vdf[rownames(vdf) %in% genes,]

for (i in 1:length(genes)){
gbar=vdf[rownames(vdf)==genes[i],]
if(length(gbar)>0){
names(gbar)=targets$Phenotype[match(names(gbar),targets$bam)]
names(gbar)= sapply(strsplit(names(gbar),split=" "),function(x) x[1])
groupcol=as.factor(names(gbar))
barplot(2^gbar,col=as.numeric(groupcol), las = 2,main=paste(genes[i]," Counts: Normalized",sep=""),ylab="normalized cpm")
}
}


```
####*Statistical Analysis of Edar group (lmFit)*

```{r}
#Run analysis of edar group:
fit1 <- lmFit(v1,design1) 
fit1 <- eBayes(fit1) 
options(digits=3) 
topgenes1=topTable(fit1,coef=3,n=50,sort.by="p")
FC = 2^(fit1$coefficients[,3])
FC = ifelse(FC<1,-1/FC,FC)
finalres=topTable(fit1,coef=3,sort="none",n=Inf)


```


####*Volcano Plot: Edar Group*


```{r results='asis', tidy=FALSE,echo=TRUE}
suppressPackageStartupMessages(library(googleVis))
library(googleVis)

op <- options(gvis.plot.tag='chart')
volcano_data=as.data.frame(cbind(fit1$coefficients[,3],-1*log10(fit1$p.value[,3])))
volcano_data$pop.html.tooltip=rownames(volcano_data)
yval=ceiling(max(fit1$coefficients[,3]))
Scatter <- gvisScatterChart(volcano_data, 
          options=list(
          tooltip="{isHtml:'True'}",
          legend='none',
          lineWidth=0, pointSize=1,
          title='Edar mut vs het', vAxis="{title:'Log Odds'}",
          hAxis="{title:'Log Fold Change'}", 
          width=1200, height=800,
          hAxes="[{viewWindowMode:'explicit', viewWindow:{min:-10, max:10}}]",
          vAxes="[{viewWindowMode:'explicit', viewWindow:{min:0, max:12}}]"))
plot(Scatter)

```

####*Simple Batch Removal Step*

Based on the linear model:

Signal = Celltype + Batch + Residual 

```{r, eval=TRUE}

#Look up Batch Effect:
fit1$design
head(fit1$coefficients)
targets
v1coeff = fit1$coefficients

#Remove batch effect from Batch2 
w1=v1
w1$E=v1$E
w1$E[,2:5]=v1$E[,2:5]-v1coeff[,2]

```

####*Heatmap: Edar Group Before and after Batch Removal*

```{r, echo=TRUE}
library(d3heatmap)
#Original Heatmap (notice batch effect)
topgenes_data=v1$E[rownames(v1$E) %in% rownames(topgenes1),]
topgenes_data=topgenes_data[match(rownames(topgenes1),rownames(topgenes_data)),]
colnames(topgenes_data)=targets$Phenotype[match(colnames(topgenes_data),targets$bam)]
colnames(topgenes_data)= sapply(strsplit(colnames(topgenes_data),split=" "),function(x) x[1])
colnames(topgenes_data)=paste(colnames(topgenes_data),targets$Batch[1:6])

d3heatmap(topgenes_data, scale = "row", dendrogram = "none", main = "title",
    color = "YlOrRd",cexRow=1,main="topgenes")

v1res=(residuals.MArrayLM(fit1,v1)) # calculate residuals
head(v1res)
v1fit=(fitted.MArrayLM(fit1,design1)) # get fitted values
head(v1fit)
head(v1$E) # original = fitted + residuals

```


```{r, echo=TRUE}
topgenes_data=w1$E[rownames(w1$E) %in% rownames(topgenes1),]
topgenes_data=topgenes_data[match(rownames(topgenes1),rownames(topgenes_data)),]
colnames(topgenes_data)=targets$Phenotype[match(colnames(topgenes_data),targets$bam)]
colnames(topgenes_data)= sapply(strsplit(colnames(topgenes_data),split=" "),function(x) x[1])
colnames(topgenes_data)=paste(colnames(topgenes_data),targets$Batch[1:6])
d3heatmap(topgenes_data, scale = "row", dendrogram = "none",
    color = "YlOrRd",cexRow=1,main="topgenes")

genes_data=v1$E[rownames(v1$E) %in% genes,]
colnames(genes_data)=targets$Phenotype[match(colnames(genes_data),targets$bam)]
colnames(genes_data)= sapply(strsplit(colnames(genes_data),split=" "),function(x) x[1])
d3heatmap(genes_data, scale = "row", dendrogram = "none", main="title",
    color = "YlOrRd",cexRow=1)

genes_data=w1$E[rownames(w1$E) %in% genes,]
colnames(genes_data)=targets$Phenotype[match(colnames(genes_data),targets$bam)]
colnames(genes_data)= sapply(strsplit(colnames(genes_data),split=" "),function(x) x[1])
d3heatmap(genes_data, scale = "row", dendrogram = "none",
    color = "YlOrRd",cexRow=1)
```

####*Barchart: Edar Group Before and after Batch Removal*

```{r, echo=TRUE}

vdf=cbind(v1$E,w1$E)
genes_count=vdf[rownames(vdf) %in% genes,]

for (i in 1:length(genes)){
gbar=vdf[rownames(vdf)==genes[i],]
if(length(gbar)>0){
names(gbar)=targets$Phenotype[match(names(gbar),targets$bam)]
names(gbar)= sapply(strsplit(names(gbar),split=" "),function(x) x[1])
names(gbar)[7:12]=names(gbar)[1:6]
groupcol=as.factor(names(gbar))
colors = c("green", "orange", "green", "violet", "orange", "red", "pink", "cyan") 
barplot(2^gbar,col=colors[as.numeric(groupcol)], las = 2,main=paste(genes[i]," Counts: Normalized",sep=""),ylab="normalized cpm")
}
}

```


####*PCA Plot: After Batch Removal*

```{r, echo=TRUE,webgl=TRUE}
edf=as.matrix(v1$E.br)
edf=as.matrix(w1$E)
tedf= t(edf)
pca=prcomp(tedf,scale.=T)
tedf1 = data.frame(tedf)

Phenotype=sapply(strsplit(targets$Phenotype[1:6],split=" "),function(x) x[1])
Batch=targets$Batch[1:6]
tedf1$group = as.factor(Phenotype)
plot(pca,type="lines")  #Decide how many PC's are relevant for plotting
pca$x[,1:3]  #look at first 3 PC's

plot3d(pca$x[,1:3],col = as.integer(tedf1$group),type="s",size=2)
group.v<-as.vector(paste(Phenotype,Batch))
text3d(pca$x, pca$y, pca$z, group.v, cex=1.0, adj = 1.2) 
#rgl.postscript("pca3d_indiv_edar_batch.pdf","pdf")

```

```{r, echo=TRUE,eval=FALSE}

finalres = cbind(w1$E,FC,finalres)
datadir="/Users/maggiec/Github/Maggie/ccbr620/Results"
setwd(datadir)
write.table(finalres,file="Gencode_FCpval_edar.txt",
      row.names=TRUE,col.names=TRUE,sep="\t",quote=FALSE)

```

###*Statistical Analysis of Fgf group*

```{r}

design2 <- model.matrix(~celltype)
fit2 <- lmFit(v2,design2) 
fit2 <- eBayes(fit2) 
options(digits=3) 
topgenes2=topTable(fit2,coef=2,n=50,sort.by="p")
finalres2=topTable(fit2,coef=2,sort="none",n=Inf)

FC2 = 2^(fit2$coefficients[,2])
FC2 = ifelse(FC2<1,-1/FC2,FC2)

```

```{r, echo=TRUE,eval=FALSE}
finalres2 = cbind(v2$E,FC2,finalres2)
datadir="/Users/maggiec/GitHub/Maggie/ccbr620/Results"
setwd(datadir)
write.table(finalres2,file="Gencode_FCpval_Fgf.txt",
      row.names=TRUE,col.names=TRUE,sep="\t",quote=FALSE)

```

####*Volcano Plot: Fgf Group*

```{r results='asis', tidy=FALSE,echo=TRUE}
suppressPackageStartupMessages(library(googleVis))
library(googleVis)

op <- options(gvis.plot.tag='chart')
volcano_data=as.data.frame(cbind(fit2$coefficients[,2],-1*log10(fit2$p.value[,2])))
volcano_data$pop.html.tooltip=rownames(volcano_data)
yval=ceiling(max(fit2$coefficients[,2]))
Scatter <- gvisScatterChart(volcano_data, 
          options=list(
          tooltip="{isHtml:'True'}",
          legend='none',
          lineWidth=0, pointSize=1,
          title='Fgf mut vs het', vAxis="{title:'Log Odds'}",
          hAxis="{title:'Log Fold Change'}", 
          width=1200, height=800,
          hAxes="[{viewWindowMode:'explicit', viewWindow:{min:-10, max:10}}]",
          vAxes="[{viewWindowMode:'explicit', viewWindow:{min:0, max:6}}]"))
plot(Scatter)

```

####*Heatmap: Fgf Group*

```{r, echo=TRUE}
library(d3heatmap)
#Original Heatmap (notice batch effect)
w2=v2
w2$E=v2$E[,c(3,5,6,1,2,4)]
topgenes_data2=w2$E[rownames(w2$E) %in% rownames(topgenes2),]
topgenes_data2=topgenes_data2[match(rownames(topgenes2),rownames(topgenes_data2)),]
colnames(topgenes_data2)=targets$Phenotype[match(colnames(topgenes_data2),targets$bam)]
colnames(topgenes_data2)= sapply(strsplit(colnames(topgenes_data2),split=" "),function(x) x[1])
colnames(topgenes_data2)=paste(colnames(topgenes_data2),targets$Batch[7:12])

d3heatmap(topgenes_data2, scale = "row", dendrogram = "none", main = "title",
    color = "YlOrRd",cexRow=1,main="topgenes")

genes_data=w2$E[rownames(w2$E) %in% genes,]
colnames(genes_data)=targets$Phenotype[match(colnames(genes_data),targets$bam)]
colnames(genes_data)= sapply(strsplit(colnames(genes_data),split=" "),function(x) x[1])
d3heatmap(genes_data, scale = "row", dendrogram = "none", main="title",
    color = "YlOrRd",cexRow=1)
```

####*Boxplots: Fgf Group*

```{r,echo=TRUE}
vdf=w2$E
for (i in 1:length(genes)){
gbar=vdf[rownames(vdf)==genes[i],]
if(length(gbar)>0){
names(gbar)=targets$Phenotype[match(names(gbar),targets$bam)]
names(gbar)= sapply(strsplit(names(gbar),split=" "),function(x) x[1])
groupcol=as.factor(names(gbar))
colors = c("blue", "yellow", "green", "violet", "orange", "red", "pink", "cyan") 
barplot(2^gbar,col=colors[as.numeric(groupcol)], las = 2,main=paste(genes[i]," Counts: Normalized",sep=""),ylab="normalized cpm")
}
else {
  paste(genes," is not present")
}
}

```


###*Provenance:*

```{r, echo=FALSE}

sessionInfo()

```